%In this work we presented a current state of the art in a field of
%self-configurable Wireless Sensor Network. 

Despite there are a lot of work in the area of self-configurable Wireless Sensor
Networks, the solutions are problem-specific, and there is no holistic approach
for developing self-adaptive WSNs. As already presented, the research has three
independent directions. Focusing on the one of the directions, researches
absolutely abandon other two, which makes it difficult to use them all together.

The development of such systems can be a long and expensive process, since the
developer should be care of all possible situations and combinations of
situations, the system can find itself in. WSNs are usually built on very
resource-restricted platforms, which makes it necessary to predefine all
possible behaviour variations, since the adaptation to unpredicted situations
could be very expensive in terms of energy and memory consumption.

Self-adaptation is also expensive in terms of performance, since the system
becomes less responsive during the analysis of the environment. There also could
be an intermediate state, when the old behaviour is not applicable in the new
situation, but the new behaviour is not chosen yet. Thus, we can consider as
costs of self-adaptivity for WSNs the reduction of both service time and
responsiveness of the system, since additional calculations consume more energy
and time. Therefore, self-adaptation for WSNs becomes an hard choice for such
systems where dynamic changes do not frequently occur. 

On the other hand, self-adaptation schemes lead to relevant advantages in terms
of performance. The most important benefit, targeted by the majority of the
works, is the energy saving. Specifically, a smart using of the resources, such
as switching off some of the nodes, reducing the transmitting distances, and so
on, as presented in Section \ref{sec:sa}, provides an useful maximization of the
battery lifetime by improving the energy consumption. 

Moreover, the self-adaptation techniques help to perform a more efficient
coverage of nodes in the standard Wireless Sensor Networks. Indeed, the
self-adaptation provides better performance by placing in an optimal way the
nodes to cover the system. At the same time, this leads to a more reliable
system by assuring the fault-tolerance in the WSNs. Interestingly, the nodes
which are not involved in the coverage process, may be activated for supplying
the interested nodes.

Finally, we can state that the most important feature introduced with the
self-adaptive wireless sensor networks is the flexibility as well as the ability
to proficiently react to the environment changes. In this sense, the
self-adaptation paradigm plays a key-role in terms of modifying system
behaviours by following the  network requirements but, at the same time,
monitoring the current performance of the system.
